\documentclass{vldb}

\usepackage{graphicx}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{hyperref}
\lstset{
  language=C,
  basicstyle=\ttfamily\scriptsize,
  tabsize=2,                    % sets default tabsize to 2 spaces
%  columns=flexible,
  xleftmargin=18pt,
  xrightmargin=1em,
  numbers=left,
  numberstyle=\scriptsize,
  numbersep=1em,
  showstringspaces=false,
  alsoletter={-},
	morekeywords={todo}, %FIXIT
	otherkeywords={TD},  %FIXIT
  captionpos=b,
  escapeinside={@}{@},
  commentstyle=\color{gray},
}


\begin{document}

\title{Incremental Memory Footprint Estimation for MonetDB}

\author{Pavlos Katsogridakis, Ying Zhang, Martin Kersten}

\maketitle

\begin{abstract}
Estimating the memory footprint of a query is an important
feature for a database system. It can answer the question whether
a query will fit in memory, and also gives the compiler the opportunity
to perform a lot of optimizations.
In this document we present Malcom (Mal ALgebra COst Model), a memory footprint
estimator, that incrementally builds a memory cost model for MAL instructions,
based on previous query executions. We evaluated our system using the tpch
benchmark suite, and a real world air traffic dataset.
\end{abstract}

\input{intro.tex}
\input{background.tex}
\input{monetdb_query_execution.tex}
\input{design.tex}
\input{experiments.tex}

\subsection*{Acknowledgments}
This research has received funding from the European Union’s Horizon 2020 research and innovation programme under Grant Agreement no. 732366 (ACTiCLOUD).

{\small
\bibliography{references}
\bibliographystyle{alpha}
}
%\printindex
\subsection*{Appendix}
In this section we collect some user scenarios. screenshots of how it should feel.
and scripts.
